---
title: "Evaluations of positive and negative stimuli using the Affective Misattribution Procedure (AMP) and self-reports"
subtitle: "Analysis"
author: "Template: Ian Hussey; content: [Student name]"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: hide
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
---

```{r, include=FALSE}

# set knit options
knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE)

# disable scientific notation
options(scipen = 999) 

```

# Dependencies

```{r}

library(tidyverse)
library(knitr)
library(kableExtra)
library(janitor)
library(scales)

```

# Data

Load the processed data and apply the global exclusions.

```{r}

data_processed <- read_csv("../data/processed/data_processed.csv")

data_processed_after_exclusions <- data_processed |>
  filter(exclude_participant == "include")

```

# Sample descriptives

## Sample size before exclusions

```{r}

data_processed |>
  count(name = "n") |>
  kable() |>
  add_header_above(header = c("Whole sample" = 1)) |> # note that you can add header rows to tables like this. The "1" indicates the number of columns the header should span. The sum of these numbers must equal the number of columns or you'll get an error.
  kable_classic(full_width = FALSE)

```

## Sample size after exclusions

Sample used in subsequent analyses

```{r}

data_processed_after_exclusions |>
  count(name = "n") |>
  kable() |>
  add_header_above(header = c("For analysis" = 1)) |>
  kable_classic(full_width = FALSE)

```

## Age

```{r}

data_processed_after_exclusions |>
  mutate(age = as.numeric(age)) |>
  summarise(Mean = mean(age, na.rm = TRUE),
            SD = sd(age, na.rm = TRUE)) |>
  mutate_all(.funs = janitor::round_half_up, digits = 1) |>
  kable() |>
  add_header_above(header = c("Age" = 2)) |>
  kable_classic(full_width = FALSE)

```

## Gender

```{r}

data_processed_after_exclusions |> 
  rename(Gender = gender) |>
  group_by(Gender) |> 
  summarise(n = n()) |> 
  mutate(Percent = paste0(round_half_up((n / sum(n)) * 100, 1), "%")) |>
  mutate(Gender = stringr::str_to_sentence(Gender)) |> # Change the case of the Gender variable so that it prints nicely
  kable() |>
  kable_classic(full_width = FALSE)

```

# Descriptives

Descriptive statistics and plots of the measures (excluding the demographics variables)

## Self-reported evaluations

### Descriptive stats

```{r}

# overall self-reported evaluations
dat_mean_ratings <- data_processed_after_exclusions |>
  summarise(Mean = mean(mean_evaluation, na.rm = TRUE),
            SD = sd(mean_evaluation, na.rm = TRUE),
            n = n()) |>
  mutate(group = "Full sample")

# self-reported evaluations by gender category
dat_mean_ratings_by_gender <- data_processed_after_exclusions |>
  group_by(group = gender) |>
  summarise(Mean = mean(mean_evaluation, na.rm = TRUE),
            SD = sd(mean_evaluation, na.rm = TRUE),
            n = n())

# combine both into one table
bind_rows(dat_mean_ratings,
          dat_mean_ratings_by_gender) |>
  select(Subset = group, Mean, SD, n) |> # select variables of interest, and rename one 
  mutate(Subset = stringr::str_to_sentence(Subset)) |> # Change the case of the Subset variable so that it prints nicely
  mutate_if(is.numeric, round_half_up, digits = 2) |>
  kable() |>
  add_header_above(header = c(" " = 1, "Self-reported evaluations" = 3)) |>
  kable_classic(full_width = FALSE)

```

### Descriptive plot

```{r}

ggplot(data_processed_after_exclusions, aes(x = mean_evaluation)) +
  geom_histogram(binwidth = 1,
                 boundary = 0,
                 fill = viridis_pal(begin = 0.45, option = "mako")(1), 
                 color = viridis_pal(begin = 0.30, option = "mako")(1)) + 
  xlab("Mean self-reported evaluation") +
  ylab("Frequency") +
  theme_linedraw() +
  scale_x_continuous(breaks = pretty_breaks(n = 7)) +
  coord_cartesian(xlim = c(1, 7)) +
  theme(panel.grid.minor = element_blank())

```

## AMP evaluations

### Descriptive stats

\TODO add table of means, SDs, Ns

```{r}


```

### Descriptive plots

```{r}

ggplot(data_processed_after_exclusions, aes(x = AMP_score)) +
  geom_histogram(binwidth = 0.05,
                 boundary = 0,
                 fill = viridis_pal(begin = 0.45, option = "mako")(1), 
                 color = viridis_pal(begin = 0.30, option = "mako")(1)) + 
  xlab("AMP score") +
  ylab("Frequency") +
  theme_linedraw() +
  scale_x_continuous(breaks = pretty_breaks(n = 10))

```

# Analyses & hypothesis tests

## Self-reported evaluations are correlated with evaluations on the AMP

### Plot

```{r}

ggplot(data_processed_after_exclusions, 
       aes(x = AMP_score,
           y = mean_evaluation)) +
  geom_jitter(color = viridis_pal(begin = 0.45, option = "mako")(1),
              alpha = 0.5) +
  geom_smooth(method = "lm",
              color = viridis_pal(begin = 0.45, option = "mako")(1)) +
  xlab("AMP score") +
  ylab("Mean self-reported evaluation") +
  theme_linedraw() 

ggplot(data_processed_after_exclusions, 
       aes(y = AMP_score,
           x = mean_evaluation)) +
  geom_jitter(color = viridis_pal(begin = 0.45, option = "mako")(1),
              alpha = 0.5) +
  geom_smooth(method = "lm",
              color = viridis_pal(begin = 0.45, option = "mako")(1)) +
  ylab("AMP score") +
  xlab("Mean self-reported evaluation") +
  theme_linedraw() 

ggplot(data_processed_after_exclusions, 
       aes(x = AMP_score,
           y = mean_evaluation)) +
  geom_jitter(color = viridis_pal(begin = 0.45, option = "mako")(1),
              alpha = 0.5) +
  xlab("AMP score") +
  ylab("Mean self-reported evaluation") +
  theme_linedraw() 

```

More complex plots:

#### Axial histograms

Scatter plots with axial histograms using ggExtra: https://cran.r-project.org/web/packages/ggExtra/vignettes/ggExtra.html

\TODO add axial histograms to a scatter plot. Split both the scatter plot and the histograms by gender.

```{r}
install.packages("ggExtra")

library("ggExtra")

install.packages("ggmarginal")
```


```{r}

# Install and load the viridis package
install.packages("viridis")
library(viridis)

# Your ggplot code with viridis color scale
p <- ggplot(data_processed_after_exclusions, aes(x = AMP_score, y = mean_evaluation, color = gender)) +
     geom_jitter(alpha = 0.5) +
     xlab("AMP score") +
     ylab("Mean self-reported evaluation") +
     theme_linedraw() +
     scale_color_viridis(discrete = TRUE)  # This will now work since viridis is loaded

# Add marginal histograms with ggExtra::ggMarginal without specifying 'fill'
p <- ggExtra::ggMarginal(p, type = "histogram", margins = "both")

# Print the plot
print(p)

```
```{r}
library(ggplot2)
library(ggExtra)
library(patchwork)

# Create a list to hold plots
plots <- list()

# Loop through each gender to create individual plots
for (gender in unique(data_processed_after_exclusions$gender)) {
  data_subset <- data_processed_after_exclusions[data_processed_after_exclusions$gender == gender,]
  
  p <- ggplot(data_subset, aes(x = AMP_score, y = mean_evaluation)) +
       geom_jitter(alpha = 0.5) +
       xlab("AMP score") +
       ylab("Mean self-reported evaluation") +
       theme_linedraw() +
       ggtitle(gender)  # Add a title to distinguish each plot
  
  # Add marginal histograms
  p <- ggExtra::ggMarginal(p, type = "histogram", margins = "both")
  
  # Store the plot in the list
  plots[[gender]] <- p
}

# Combine the plots using patchwork
combined_plot <- wrap_plots(plots)

# Print the combined plot
print(combined_plot)

```

#### Labelled points

Label points using ggrepel: https://cran.r-project.org/web/packages/ggrepel/vignettes/ggrepel.html

\TODO Label the points in a scatter plot using their participant codes.

```{r}
install.packages("ggrepel")
library(ggrepel)
```
```{r}
library(ggplot2)
library(ggrepel)

# Define the 'label' column with criteria for significant points
# Here I assume `AMP_score` and `mean_evaluation` are your criteria for significance
data_processed_after_exclusions$label <- ifelse(
  data_processed_after_exclusions$AMP_score > 0.8 | data_processed_after_exclusions$mean_evaluation > 4.5,
  as.character(data_processed_after_exclusions$participant_code),  # Convert factors to characters if necessary
  NA
)

# Create the ggplot object 'p' with labels for participant codes
p <- ggplot(data_processed_after_exclusions, aes(x = AMP_score, y = mean_evaluation)) +
     geom_jitter(aes(color = gender), alpha = 0.5) +
     geom_text_repel(aes(label = subject), na.rm = TRUE) +  # Label points using 'subject'
     xlab("AMP score") +
     ylab("Mean self-reported evaluation") +
     theme_minimal()

# Print the plot
print(p)


```

#### Magnify areas

Magnify areas of your plot with ggmagnify: https://hughjonesd.github.io/ggmagnify/

\TODO Magnify an area of one of your scatter plots, eg where there are a lot of data points in a small area.

```{r}
install.packages("ggmagnify", repos = c("https://hughjonesd.r-universe.dev", 
                 "https://cloud.r-project.org"))
```

```{r}
library(ggplot2)
library(ggmagnify)

# Create the ggplot object 'p'
p <- ggplot(data_processed_after_exclusions, aes(x = AMP_score, y = mean_evaluation, color = gender)) +
     geom_jitter(alpha = 0.5) +
     xlab("AMP score") +
     ylab("Mean self-reported evaluation") +
     theme_minimal()

# Define the source area to magnify
from <- c(xmin = 0.3, xmax = 0.6, ymin = 0.7, ymax = 1.3)

# Define the destination area for the magnified plot (adjust these values as needed)
to <- c(xmin = 0.75, xmax = 1, ymin = 4, ymax =5 )

# Add the magnified area to the plot
p + geom_magnify(from = from, to = to)

```

### Test

\TODO 

## Self-reported evalautions differ between men and women

### Plot

\TODO split histogram, split violin plot, raincloud plot, etc.


```{r}
  # Install and load required packages
  if (!require('ggplot2')) {
    install.packages("ggplot2")
  }
  library(ggplot2)
  
  if (!require('ggrepel')) {
    install.packages("ggrepel")
  }
  library(ggrepel)
  
  # Define the height for raincloud plot components
  rain_height <- .1
  
  # Create a raincloud plot using ggplot2 and ggrepel
  p_raincloud <- ggplot(data_processed_after_exclusions, aes(x = gender, y = mean_evaluation, fill = gender)) +
    # clouds (flat violins)
    geom_violin(trim = FALSE, alpha = 0.4, scale = "width",
                position = position_nudge(x = rain_height + .05)) +
    # rain (points)
    geom_point(aes(color = gender), size = 2, alpha = 0.5, position = position_jitter(width = rain_height)) +
    # boxplots
    geom_boxplot(width = rain_height, alpha = 0.4, outlier.shape = NA,
                 position = position_nudge(x = -rain_height * 2)) +
    # summary statistics (mean and standard error)
    stat_summary(fun.data = mean_se, geom = "errorbar", width = rain_height, 
                 position = position_nudge(x = rain_height * 3), color = "black") +
    stat_summary(fun = mean, geom = "point", 
                 position = position_nudge(x = rain_height * 3), color = "black", size = 3) +
    # adjust layout
    scale_x_discrete(name = "", expand = expansion(add = c(rain_height * 3, 0))) +
    scale_y_continuous(name = "Mean Self-Reported Evaluation") +
    coord_flip() +
    facet_wrap(~gender, scales = "free_y", ncol = 1) +
    # custom colors and theme
    theme_minimal() +
    theme(legend.position = "none")  # Hide the legend if not needed
  
  # Print the plot
  print(p_raincloud)

```

### Test

\TODO 

## Evaluations on the Affect Misattribution Procedure differ between men and women

### Plot

\TODO split histogram, split violin plot, raincloud plot, etc.

This time, vary the labeling and order of the legend, e.g., capitalise "Men" and "Women", and know how to change the order of the factors.

```{r}



```

### Test

\TODO 

```{r}



```

## Combining plots

Combine plots using patchwork: https://patchwork.data-imaginist.com/

\TODO Combine at least three of the above plots into one.

```{r}



```

## Saving plots

Save plots to disk with `ggsave()`

\TODO Save the above combined plot to disk as both .png and .pdf. Ensure the png has at least 300dpi resolution.

```{r}



```

# Session info

```{r}

sessionInfo()

```



